---
title: "IST 777 practice exam #4"
author: "Abhijith Anil Vamadev"
output: pdf_document
---

# 1.	Help

_Run ?USJudgeRatings and examine the help file for that built-in data frame. Cut and paste the most relevant information from the help screen below._

```{r}
USJudgeRatings
?UsJudgeRatings
#Lawyers' ratings of state judges in the US Superior Court.
```


# 2.	New data frame

_Copy the built-in data frame USJudgeRatings into a new data frame. Call your new data frame “newJudge”._

```{r}
newJudge <- USJudgeRatings
```


# 3.	Binary variable

_Create a binary version of the RTEN variable in newJudge with the following code:_

```{r }
newJudge$retain <- newJudge$RTEN > median(newJudge$RTEN)
```

# 4.	Exploratory analysis

_Carry out the necessary univariate and bivariate explorations of the variables retain, CONT, and INTG from your newJudge data frame._
```{r}
hist(newJudge$RTEN)
hist(newJudge$CONT)
hist(newJudge$INTG)
```



# 5.	Logistic regression

_Run a logistic regression model using CONT and INTG to predict retain. Be sure to check the regression diagnostics._


# 6.	Writeup

_Provide a write up of your results, with an interpretation of the research question, “Can the number of contacts between a lawyer and a judge along with the lawyer’s rating of judicial integrity predict retention recommendations about the judge?”_


