---
title: 'Week 3 – Sampling Exercise #1'
output:
  pdf_document: default
  html_notebook: default
---

Instructions: In the asynchronous material for this week, we created and made sense of sampling distributions by creating a simulated population from random data and then drawing samples from that population. By examining large collections of samples, we can understand what to expect when we make inferences from a single sample of real data. In this first live exercise, we’ve used a few lines of code to set up another simulated population: scores on a test of achievement. Achievement tests like this one are calibrated so that the population mean is 100 and the population standard deviation is 10. The diagrams below depict a simulated population of N=100,000 test takers, each of whom scored somewhere between 60 and 140 on the test. Using the diagrams, answer the following questions about the simulated population.

```{r}
set.seed(1234)                                        # Control randomization
testPop <- rnorm(100000, mean=100, sd=10)             # Create simulated population 
testPopMean <- mean(testPop)                          # Calculate mean
sumSq <- sum((testPop-testPopMean)^2)                 # Sum of squares
testPopVar <- sumSq/length(testPop)                   # Variance
testPopSD <- sqrt(testPopVar)                         # Standard deviation
```

1. Here’s a histogram of the population. What is the shape of this distribution and the formal statistical name for that shape:

```{r}
hist(testPop)
```

2. In this next histogram, I’ve marked the 1st, 2nd, and 3rd quartiles. Define the term quartile and estimate the values associated with these quartiles. The second quartile (also known as the 50th percentile) has a special name. What is it?

```{r}
# Markings of quartiles
hist(testPop)                                           
abline(v=quantile(testPop, probs=0.25),col="orange")    # First quartile
abline(v=quantile(testPop, probs=0.5),col="orange")     # Second quartile
abline(v=quantile(testPop, probs=0.75),col="orange")    # Third quartile
```


3. The 1st and 3rd quartiles in that histogram are very close to the middle. Can that be right? What percentage of cases fall below the 1st quartile? Above the 3rd quartile? If you want, you can write a few lines of code to confim the answer. 


4. By definition, this achievement test has a population standard deviation of 10. I’ve used abline() to mark points that are one and two standard deviations below the mean. Examine the histogram below and report what value is two standard deviations below the mean and what value is two standard deviations above the mean:

```{r}
# Markings of one and two standard deviations 
hist(testPop)
abline(v=90, col="red")
abline(v=80, col="blue")
abline(v=110, col="red") 
abline(v=120, col="blue")
```

5. Now I’ve marked some different quantiles, specifically the 0.025 quantile and the 0.975 quantile. Estimate where these quantiles fall on the X-axis. What percentage of cases occur between the 0.025 quantile and the 0.975 quantile?

```{r}
# Markings of the central region and tails 
hist(testPop)
abline(v=quantile(testPop, probs=0.025),col="green") 
abline(v=quantile(testPop, probs=0.975),col="green")
```

6. The area between the green lines is known as the “central region” and the two areas outside of the green lines are the “tails.” What percentage of cases fall in the lower tail, that is, below the 0.025 quantile in the histogram above? What percentage of cases fall in the upper tail, that is, above the 0.975 quantile in the histogram above?

