---
title: "IST772 Problem Set 11"
author: "Abhijith Anil Vamadev"
output:
  pdf_document: default
---

The homework for week 12 is based on exercises 2, 5, 6, 7, and 8 on pages 272 and 273 but with changes as noted in this notebook (i.e., follow the problems as given in this document and not the textbook). 

Attribution statement: (choose only one)
1. I did this homework by myself, with help from the book and the professor

# Chapter 11, Exercise 1

_Download and library the nlme package and use data("Pixel") to activate the Pixel data set and "? Pixel" for documentation. Inspect the data and create a box plot showing the Pixel intensities on different days. (1 pt) Run a repeated measures ANOVA to compare pixel intensities on days 4, 6, 10 and 14 using aov() (1 pt) and report the results. (1 pt) You can use a command like:_

```{r}
#install.packages('nlme')
library(nlme)
myData <-subset(Pixel, day %in% c(4,6,10,14))
boxplot(pixel~day, myData)
```

_to subset the data. Keeping in mind that the data will need to be balanced before you can conduct this analysis. (1 pt) Try running a command like:_

```{r}
new <- myData
table(myData$Dog,myData$day)
new$dayFact <- as.factor(new$day) # Convert Time to a factor
list <- rowSums(table(new$Dog,new$dayFact))==8# Make a list of rows
list <- list[list==TRUE] # Keep only those with 8 observations
list <- as.numeric(names(list)) # Extract the row indices
new <- new[new$Dog %in% list,] # Match against the data
table(new$Dog,new$dayFact) #result

aovOut<-aov(pixel ~ day+Error(Dog), data = new)
summary(aovOut)
```

_as the starting point for cleaning up the data set._


* In the first section of the output, marked “Error: Dog,” the “Residuals”
refers to the variance attributable to individual differences among chicks. The
df = 7 signifies that we have 8 Dogs in the data set. The sum of squares of
48,506 represents variation in the weight variable that is directly attributable
to individual differences among dogs.
* In the second section, marked “Error: Within,” the effect of Time is
expressed as an F-ratio, F(11,484) = 231.6,p < .001. The df = 1 for the numerator 
reflects the 1 points in time where we measured the weights for each of
the 7 chicks. The df = 5 for the denominator is the remaining error variance that is 
not attributable to individual differences. This F-ratio tests the null hypothesis that
changes in Dog pixels are consistently 0 across all time intervals. With a p-value of 
0.0618 that is not below the conventional threshold of significance
we fail to reject this null hypothesis that changes in Dog pixels are consistently 0 across all time intervals.


# Chapter 11, Exercise 5

_Given that the built-in AirPassengers data set (see "? AirPassengers" for documentation) has a substantial growth trend, use diff() to create a differenced data set. (1 pt) Use plot() to examine and interpret the results of differencing. Use cpt.var() to find the change point in the variability of the differenced time series. (1 pt) Plot the result and describe in your own words what the change point signifies. (1 pt)_
```{r}
AirPassengers
library(changepoint)
difference <- diff(AirPassengers)
plot(difference)
plot(cpt.var(difference))

```
* As the years grew, there is a definite difference increase among air passengers. More monthly totals of international airline passengers are visible through the years. 
* The redline in the plot indicates the changepoint of the dataset. The changepoint scans the data and find a major shift occured in the mean level of the data. In this plot, the major shift occured in point 76, which is around mid 1955 where the amount of international passengers increased by about 45,000 in one month.

# Chapter 11, Exercise 6

_Use cpt.mean() on the undifferenced AirPassengers time series. (1 pt) Plot and interpret the results. Compare the change point of the mean that you uncovered in this case to the change point in the variance that you uncovered in Exercise 5. What do these change points suggest about the history of air travel? (1 pt)_
```{r}
plot(cpt.mean(AirPassengers, class = FALSE))
```
* These change points indicates that there has been a shift in the mean over time.

# Chapter 11, Exercise 7

_Find historical information about air travel on the Internet and/or in reference materials that sheds light on the results from Exercises 5 and 6. Write a mini-article (less than 250 words) that interprets your statistical findings from Exercises 5 and 6 in the context of the historical information you found. (1 pt)_
```{r}
#Source:https://www.loveexploring.com/gallerylist/86315/how-air-travel-
#has-changed-in-every-decade-from-the-1920s-to-today
```

* In 1920s, air travel was mostly viewed as a luxury item which was only affordable by the very upper class in socity, and travelers were all rich members from the society. Passengers were treated similarly; they normally find food and drink. However, during 1950s, this trend began to shift. Air travel became more like a necessity. Big airplane started to take a train of passengers. This might explain the results in question 5 and 6. There was a significant passenger increase in 1955.

# Chapter 11, Exercise 8

_Use bcp() on the AirPassengers time series. (1 pt) Plot and interpret the results. Make sure to contrast these results with those from Exercise 6. (1 pt)_
```{r}
library(bcp)
bcpAD <- bcp(as.vector(difference))
plot(bcpAD)
```

* The upper pane shows the original
time series and the lower pane shows the probabilities of a mean change at each
point in time. You will note that there are isolated spikes that show probabilities near 1 
at many points across the timeline. Yet, somewhere after the data point of 100, more spikes 
can be observed, we can see there is
substantial density of probability values near 1

* The posterior means and probability plots showed a shift in probability and mean
difference around the same location in question 6, 76 was the change point in the 
previous result, so it can be explained that there was a big shift occurred in the data at this point and after.

